---
title: "Data Visualization"
author: "Nicole Doyle"
date: "8/21/2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Data Visualization
####### R Markdown
  use to adjust style to text 

### 3.1 Introduction
Tidyverse is a library full of data sets and help pages, and functions that the code can access. 
```{r}
library(tidyverse)
# we must load the tidyverse every session
```

### 3.2 First Steps 

QUESTION: Do cars with big engines use more gas than cars with small engines?

##### 3.2.1 The mpg Data Frame

A data frame is a rectangular collection of variables (in the columns) and observations (in the rows). mpg contains observations collected by the US Environmental Protection Agency on 38 models of car.

Use  ?mpg to access the help pages for mpg
```{r}
mpg
```
##### 3.2.2 Creating A ggplot

```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy))

#ggplot(data = mpg) specifically accesses the mpg dataset from the ggplot package.Then, add layers through geom_point, which is a scatter plot, geom_point(mapping = aes(x = displ, y = hwy) referer to the x and y values are going to be mapped on the graph. 
```

This plot shows the negative relationship between engine size (displ) and the fuel efficiency (hwy). Larger engine = more fuel


##### 3.2.4  Exercises

> 1. Run ggplot(data = mpg). What do you see?

  Background, Canvas for graph, so layers of data poinnts go on top. This code creates an empty plot. The ggplot() function creates the background of the plot, but since no layers were specified with geom function, nothing is drawn.
```{r}
ggplot(data = mpg)
```

> 2. How many rows are in mpg? How many columns?

 234 rows and 11 columns in the mpg data frame

> 3. What does the drv variable describe? Read the help for ?mpg to find out. 

  The drv variable is a categorical variable which categorizes cars into front-wheels, rear-wheels, or four-wheel drive. 
```{r}
?mpg
```

> 4. Make a scatterplot of hwy vs cyl.

```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = cyl, y = hwy))
```

> 5. What happens if you make a scatterplot of class vs drv? Why is the plot not useful? 

  A scatter plot is not useful in this case because both variables are CATEGORICAL, it doesn't show us how many observations there are of each variable. Scatter plots are better suited for plotting continuous variables. 
```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = class, y = drv))
```

### 3.3 Aesthetic Mappings

Here we are adding a third variable to the graph by mapping it to an aesthetic, which is a visual property such as size shape or color. In this case you would attribute an aesthetic to another variable and put it in the argument of aes with the other variables.

ggplot2 will automatically assign a unique level of the aesthetic (ex. an unique color) to each unique value of the variable, a process known as scaling. ggplot2 will also add a legend that explains which levels correspond to which values. 

```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy, color = class))
```

In the example above, we mapped class to the color aesthetic, but we could have mapped class to the size aesthetic in the same way. In this case, the exact size of each point would reveal its class affiliation. We get a warning here because mapping an unordered variable (class) to an ordered aesthetic (size) is not a good idea.
```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy, size = class))
# Not as good for this case because the numbers are so close together/small
```

Or we could have mapped class to the alpha aesthetic which controls the transparency of the points, or to the shape aesthetic, which controls the shape of the points. 
See alpha: 
```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = displ,  y = hwy, alpha = class))
#alpha refers to the transparency of the points, again not very good for this data
```

What happened to the SUVs? ggplot2 will onlyi use six shapes at a time. By default, additional groups will go unplotted when you use the shape aesthetic. 
See shape:
```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy, shape = class))
#too many different values because there are only 6 different shapes. 
```

Something to note is that if you just want to change the aesthetics of the graph such as the color of every point and not representing a variable the code that dictates the color should be outside the argument of aes because it is no longer a variable.

For each aesthetic, you use aes() to associate the name of the aesthetic with a variable to display. The aes() function gathers together each of the aesthtic mappings used by a layer and passes them to the layer's mapping argument. The syntax highlights a useful insight about x and y: the x and y locations of a point are themselves aesthetics, visual properties that you can map to variables to display information about the data. 

Once you map an aesthetic, ggplot2 takes care of the rest. It selects a reasonable scale to use with the aesthetic, and it constructs a legend that explains the mapping between levels and values. For x and y aesthetics, ggplot2 does not create a legend, but it creates an axis line with tick marks and a label. The axis line acts as a legnd; it explains the mapping between locations and values.

You can also set the aesthetic properties of your geom manually. For example, we can make all of the points in our plot blue:
```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy), color = "blue")
#make sure to have the code for blue outside the variable argument. 
```


##### 3.3.1 Exercises

> 1. What’s gone wrong with this code? Why are the points not blue?

```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy, color = "blue"))
```

The color = "blue" is in the variable argument so the R creates another variable "blue" instead of coloring the data points. 

RC: The argument color = "blue" is included within the mapping argument, and as such, it is treated as an aesthetic, which is a mapping between variables and a value. 

The following code does produce the expected result.

```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy), color = "blue")
```


> 2. Which variables in mpg are categorical? Which variables are continuous? (Hint: type ?mpg to read the documentation for the dataset). How can you see this information when you run mpg?

```{r}
?mpg
```
 displ, year, cyl, cty, hwy are continuous and the manufacturer, model, trans, fl, and class, drv are all categorical. 

> 3. Map a continuous variable to color, size, and shape. How do these aesthetics behave differently for categorical vs. continuous variables?

```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy, color = cty))
```

```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy, size = cty))
```

```{r}
#ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy, shape = cty))
```

```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy, color = class))
```


> 4. What happens if you map the same variable to multiple aesthetics?

It is redundant because you it just represents the same information twice on the same graph. 

```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy, color = hwy))
```


> 5. What does the stroke aesthetic do? What shapes does it work with? (Hint: use ?geom_point)

The ggplot() function behaves as if a temporary variable was added to the data with values equal to the result of the expression. In this case, the result of displ < 5 is a logical variable which takes values of TRUE or FALSE. 

> 6. What happens if you map an aesthetic to something other than a variable name, like aes(colour = displ < 5)? Note, you’ll also need to specify x and y.

Aesthetics can also be mapped to expressions like displ.

```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy, colour = displ < 5))
```


### 3.5 Facets

Facets are another way to add variables, especially categorical variable. It splits the plot into different facets, or subplots, that each display one subset of data. To facet the plot by a single variable use facet_wrap(), then put ~ and the variable name.  To facet a plot on the combination of two variables, use facet_grid() with the two variable names split with ~. If you don't want to facet in the rows or columns use . instead of a variable name respectively. 

```{r}
ggplot(data=mpg) + geom_point(mapping = aes(x = displ, y = hwy)) + facet_wrap(~ class, nrow = 2)
#facets are a kind of mini graph, facets are for categorical variables
```

To facet your plot on the combination of two variables, add facet_grid() to your plot call. The first argument of facet_grid() is also a formula. Tis time the formula should contain two variable names separated by a ~.
```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy)) + facet_grid(drv ~ cyl)
# just note how drv is first but is on the y-axis and cyl is on the x axis even though we usually think of (x,y)
```

If you prefer to not facet in the rows or columns dimension, use a . instead of a variable name, e.g. + facet_grid(. ~ cyl).

##### 3.5.1 Exercises

> 1. What happens if you facet on a continuous variable?

The continuous variable is converted to a categorical variable, and the plot contains a facet for each distinct value.
```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy)) + facet_wrap(~ cty, nrow = 1)
```

> 2. What do the empty cells in plot with facet_grid(drv ~ cyl) mean? How do they relate to this plot?

There's empty cells because there are no data points that correspond to the both conditions. It relates to the plot below because the plot is drv and cyl and the points where there is no point on the plot is where there are empty cells in the facet graph. 

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = drv, y = cyl))
```

> 3. What plots does the following code make? What does . do?

It takes the place of another variable and put the other variable on the desired axis.
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(drv ~ .)
```

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(. ~ cyl)
```

> 4. Take the first faceted plot in this section:

```{r}
ggplot(data = mpg) +  geom_point(mapping = aes(x = displ, y = hwy)) + facet_wrap(~ class, nrow = 2)
```

What are the advantages to using faceting instead of the color aesthetic? What are the disadvantages? How might the balance change if you had a larger dataset?

You can more easily pick out data if you are looking for a specific category of data. But its harder to see the individual points in the facets and the overall trends of the graph based on the continuous values. If there was a larger data set it might be harder to see all the individual points but also easier if the data was split into facets to look at individually. 

Advantages of encoding class with facets instead of color include the ability to encode more distinct categoires. Facets allow more readability than just a slight difference in colors. Given human visual perception, the max number of colors to use when encoding unordered categorical (qualitative) data is nine, and in practice, often much less than that. Displayhing ovservations from different categories on different scales makes it difficult to directly compare  values of observations across categories. However it can make it easier to compare the shape of the relationship between the x and y variables across categories.

Disadvantages of encoding the clas variable with facets instead of the color aesthetic include the difficulty of comparing the values of observations between categories since the observations for each category are on different plots. Using the same x- and y-scales for all facets makes it easier to compare vales of observations across categories, but it is still more difficult than if the had been displayed on the same plot. Since encoding class within color also places all points on the same plot, it visualizes the unconditional relationship between the x and y variables; with facets, the unconditional relationship is no longer visualized since the poinst are spread across multiple plots. 
The benefit of encoding a variable with faceting over encoding it with color increase in both the number of poitns and number of categories. With a large number of points, there is often overlap. It is difficult to handle overlapping points with different colors color. Jittering will still work with color. But jittering will only work well if there are a few points and the classes do not overlap much, otherwise, the colors of areas will no longer be distinct, and it will be hard to pick out th patterns of different categories visually.

Transparency (alpha) does not work well with colors since the mixing of overlapping transparent colors will no longer represent the colors of the categories. Binning methods already use color to encode the density of points in the bin, so color cannot be used to encode categories.
  As the numbers of categories increase, the difference between colors decreases, to the point that the color of categories will no longer be visually distinct. 


> 5. Read ?facet_wrap. What does nrow do? What does ncol do? What other options control the layout of the individual panels? Why doesn’t facet_grid() have nrow and ncol arguments?

nrow dictates the number of rows and ncol the number of columns. scales, shrink, dir, and switch also control the layout. facet_grid doesn't use nrow and ncol beause the number of variables inside it determines the rows and columns. 

RC: The arguments nrow (ncol) determines the number of rows (columns) to use when laying out the facets. It is necessary since facet_wrap() only facets on one variable.

The nrow and ncol arguments are unnecessary for facet_grid() since the number of unique values of the variables specified in the function determines the number of row and columns.

```{r}
?facet_wrap
```

> 6. When using facet_grid() you should usually put the variable with more unique levels in the columns. Why

On a computer scree you will have more space on horizontally than vertically.
```{r}
ggplot(data = mpg) +geom_point(mapping = aes(x = displ, y = hwy)) + facet_grid(cty ~ cyl)
```

The columns are on the longer side of the rectangle, so there is more space for the more unique values. 

### 3.6 Geometric Objects 

A geom is the geometrical object that a plot uses to represent data. People often describe plots by the type of geom that the plot uses. For example, bar charts use bar geoms, line charts use line geoms, boxplots use boxplot geoms, and so on. Scatterplots break the trend; they use the point geom. As we see above, you can use different geoms to plot the same data.

To chnge the geom in your plot, change the geom function that you add to ggplot(). For example:

```{r}
ggplot(data = mpg) + geom_smooth(mapping = aes(x = displ, y = hwy))
#using geom_smooth, it graphs the same points, but give it a line of fit, the gray represents the error
```

Every geom function in ggplot2 takes a mapping argument. However, not every aesthetic works with every geom. You could set the shape of a point, but you couldn't set the "shape" of a line. On the other hand, you could set the linetype of a line. geom_smooth() will draw a different line, with a different linetype, for each unique value of the variable that you map to linetype. 
```{r}
ggplot(data = mpg) + geom_smooth(mapping = aes(x = displ, y = hwy, linetype = drv))
#linetype divided the smooth graph by the variable drv
```

Here geom_smooth() separates the cars into three lines based on their drv value, which describes a car's drivetrain. One line describes all of the points with a 4 value, one line describes all of the points with an f value, and one line describes all of the points with an r value. Here, 4 stands for four-wheel drive, f for front-wheel drive, and r for rear-wheel drive.

if this sounds strange, we can make it more clear by overlaying the lines on top of the raw data and then coloring everything according to drv. 
```{r}
ggplot(data = mpg) + geom_smooth(mapping = aes(x = displ, y = hwy, group = drv))
#divides the smooth graph by grouping into different drv, there is no legend when using group
```

Notice that this plot contains two geoms in the same graph!

ggplot2 provides over 40 geoms, and extension packages provide even more (see https://exts.ggplot2.tidyverse.org/gallery/ for a sampling). The best way to get a comprehensive overview is the ggplot2 cheatsheet, which you can find at http://rstudio.com/cheatsheets. 

Many geoms, like geom_smooth(), use a single geometric object to display multiple rows of data. For these geoms, you can set the group aesthetic to a categorical variable to draw multiple objects. ggplot2 will draw a separate object for each unique value of the grouping variable. In practice, ggplot2 will automatically group the data for these geoms whenever you map an aesthetic to a discrete variable (as in the linetype example). It is convenient to rely on this feature because the group aesthetic by itself does not add a legend or distinguishing features to the geoms.
```{r}
ggplot(data = mpg) +geom_smooth(mapping = aes(x = displ, y = hwy, color = drv))
#splits the smooth graph into drv by color
```

To display multiple geoms in the same plot, add multiple geom functions to ggplot():

```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy)) + geom_smooth(mapping = aes(x = displ, y = hwy))
# just add another layer to combine both smooth and point 
```

This, however, introduces some duplication in our code. Imagine if you wanted to change the y-axis to display cty instead of hwy. You’d need to change the variable in two places, and you might forget to update one. You can avoid this type of repetition by passing a set of mappings to ggplot(). ggplot2 will treat these mappings as global mappings that apply to each geom in the graph. In other words, this code will produce the same plot as the previous code:

```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy, color = drv)) + geom_smooth(mapping = aes(x = displ, y = hwy, color = drv))
```

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = drv)) + geom_point() + geom_smooth() 
#condense the need for repeating the variables by having them inside then adding the geom layers after
```
If you place mappings in a geom function, ggplot2 will treat them as local mappings for the layer. It will use these mappings to extend or overwrite the global mappings for that layer only. This makes it possible to display different aesthetics in different layers.
```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + geom_smooth() + geom_point(mapping = aes(color = class))
```

###### 3.6. Exercises 1-5

> 1.What geom would you use to draw a line chart? A boxplot? A histogram? An area chart?

geom_line = line chart
geom_boxplot = boxplot
geom_histogram = hiostogram
geom_area = area chart


>2. Run this code in your head and predict what the output will look like. Then, run the code in R and check your predictions.

This code produces a scatter plot with displ on the x-axis, hwy on the y-axis, and the points colored by drv. There will be a smooth line, without standard errors, fit through each drv group.

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = drv)) + 
  geom_point() + 
  geom_smooth(se = FALSE) 
#(se = FALSE) takes out standard error indicator on the graph
```

> 3.What does show.legend = FALSE do? What happens if you remove it?
Why do you think I used it earlier in the chapter?

The theme option show.legend = FALSE hides the legend box.

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = drv)) + geom_point(show.legend = FALSE) + geom_smooth(show.legend = FALSE)
```


show.legend = FALSE makes it so the legend doesn't show up on the graph. 


> 4. What does the se argument to geom_smooth() do?

It adds standard error bands to the lines.
> 5. Will these two graphs look different? Why/why not?

No its the same.

RC: No. Because both geom_point() and geom_smooth() will use the same data and mappings. They will inherit those options from the ggplot() object, so the mappings don’t need to specified again.
```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point() + 
  geom_smooth()
```

```{r}
ggplot() + 
  geom_point(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_smooth(data = mpg, mapping = aes(x = displ, y = hwy))
```

### 3.7 Statistical transformations

```{r}
diamonds
```

```{r}
ggplot(data = diamonds) + geom_bar(mapping = aes(x = cut))
```

### 3.8 Position adjustments

```{r}
ggplot(data = diamonds) + geom_bar(mapping = aes(x = cut, color = cut))
```

```{r}
ggplot(data = diamonds) + geom_bar(mapping = aes(x = cut, fill = cut))
```

```{r}
ggplot(data = diamonds) + geom_bar(mapping = aes(x = cut, fill = clarity))
```

```{r}
ggplot(data = diamonds, mapping = aes(x = cut, fill = clarity)) + geom_bar(alpha = 0.2, position = "identity")
```

```{r}
ggplot(data = diamonds, mapping = aes(x = cut, color = clarity)) + geom_bar(fill = NA, position = "identity")
```

```{r}
ggplot(data = diamonds, mapping = aes(x = cut, fill  = clarity)) + geom_bar(position = "fill")
```

```{r}
ggplot(data = diamonds, mapping = aes(x = cut, fill = clarity)) + geom_bar(position = "dodge")
```

```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy))
```

```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy), position = "jitter")
```

##### 3.8.1 Exercises

> 1. What is the problem with this plot? How could you improve it?

The graph does not display all the data points in the data set. Jittering the graph could improve it by showing all the points on the graph. 

```{r}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) + 
  geom_point() 

```

> 2. What parameters to geom_jitter() control the amount of jittering?

width and height control the amount of jittering. 
```{r}
?geom_jitter()
```
 
> 3. Compare and contrast geom_jitter() with geom_count().

There are similar size dots in the same places indicating then number of overlapping points, but in the non jittered verson the points congregate in vertical line and in the jittered version they are not aligned on those vertical lines as much becaue they are jittered to the side.

```{r}
ggplot(data = mpg) + geom_count(mapping = aes(x = displ, y = hwy))
```

```{r}
ggplot(data = mpg) + geom_count(mapping = aes(x = displ, y = hwy), position = "jitter")
```

> 4. What’s the default position adjustment for geom_boxplot()? Create a visualisation of the mpg dataset that demonstrates it. 

The default position of geom_boxplot() is dodge2.

```{r}
?geom_boxplot
```

```{r}
ggplot(data = mpg) + geom_boxplot(mapping = aes(x = hwy))

```















