---
title: "exploring"
author: "Nicole Doyle"
date: "10/30/20"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load the necessary libraries
```{r}
library(tidyverse)
library(readr)
listings <- read_csv("C:/Users/ndai7/Desktop/Git Directories/exploring-ndoyle3146/listings.csv")
View(listings)
```

Take a look inside your dataset
```{r}
listings
```

### Variation

Perform an analysis of the variation in the "neighbourhood" column. 

```{r}
ggplot(data = listings) + 
  geom_bar(mapping = aes(x = neighbourhood)) + 
  theme(axis.text.x = element_text(angle = 90))

```

* Which values are the most common? Why?

The Buttes-Montmartre is the most common neighborhood. This neighborhood could be the most common because it contains tourist locations like the Basilica of the Sacré-Cœur and is also the nightclub district. This could make it more popular for air bnbs.

* Which values are rare? Why? Does that match your expectations?

The Louvre neighborhood is the least common neighborhood, this did not match my expectations because I thought the location would be good for visiting the Louvre, but this could also be a reason why there is less airbnbs there, because of the high numbers of tourists. 

* Can you see any unusual patterns? What might explain them?

I don't really ses any unusual patterns. 

Perform an analysis of the variation in the "room_type" column. 

```{r}
ggplot(data = listings) + 
  geom_bar(mapping = aes(x = room_type))
```

* Which values are the most common? Why?

Most commonly the airbnb has the entire home/apt available. This makes sense because people would prefer to have the house to themselves rather than shared or just one room. 

* Which values are rare? Why? Does that match your expectations?

The shared room airbnbs are the least common. This makes sense because toursits or people who come into the city would most likely not want to share a room with another person. 

* Can you see any unusual patterns? What might explain them?

I don't see any unusual patterns. 

Perform an analysis of the variation in the "price" column. Make sure to explore different "binwidth" values in your analysis.

```{r}
ggplot(data = listings) + 
  geom_histogram(mapping = aes(x = price))

ggplot(data = listings) +
  geom_histogram(mapping = aes(x = price), binwidth = 100) +
   coord_cartesian(xlim = c(0, 1750))

ggplot(data = listings) +
  geom_histogram(mapping = aes(x = price), binwidth = 30) +
  coord_cartesian(xlim = c(0, 1750))

```

* Which values are the most common? Why?

There is a spike in values that are cheaper, but there is lower amounts of value closest to 0. This could be because the cheapest airbnbs could be the ones that less people want because of lower quality, but then people also want the cheapest good quality airbnbs. 

* Which values are rare? Why? Does that match your expectations?

As the prices get over about 300 there is not many airbnbs. This makes sense because these would be either very nice places, which is more rare, or just very expensive, which less people would want. 

* Can you see any unusual patterns? What might explain them?

I don't see any unusual patterns. 


Perform an analysis of the variation in the "minimum_nights" column. Make sure to explore different "binwidth" values in your analysis.

```{r}
ggplot(data = listings) +
  geom_histogram(mapping = aes(x = minimum_nights), binwidth = 20) +
   coord_cartesian(xlim = c(0, 200))

ggplot(data = listings) +
  geom_histogram(mapping = aes(x = minimum_nights), binwidth = 10) + 
  coord_cartesian(xlim = c (0, 200))
```

* Which values are the most common? Why?

The lowest values are the most common because people who are visiting for a short time would like the airbnbs where the minimum nights is very low

* Which values are rare? Why? Does that match your expectations?

As the minimum_nights gets higher there are less airbnbs. This makes sense because customers are more likely to only want to rent for a few nights so that would be the most common type of airbnb. 

* Can you see any unusual patterns? What might explain them?

There is a slight increase around the 100 day minimum nights. This might be because if customers are renting the airbnb for a entire summer. 

Perform an analysis of the variation in the "number_of_reviews" column. Make sure to explore different "binwidth" values in your analysis.

```{r}
ggplot(data = listings) + 
  geom_histogram(mapping = aes(x = number_of_reviews), binwidth = 20) + 
  coord_cartesian(xlim = c(0, 400))

ggplot(data = listings) +
  geom_histogram(mapping = aes(x = number_of_reviews), bindwidth = 50) + 
  coord_cartesian(xlim = c(0, 400))
```

* Which values are the most common? Why?

There are many airbnbs that do not have any reviews. This could because there are so many airbnbs that theres is likely to be many where people don't leave reviews.

* Which values are rare? Why? Does that match your expectations?

Airbnbs with over 100 reviews are very rare. The count of airbnbs gets more rare as the number of reviews increases.

* Can you see any unusual patterns? What might explain them?
There are not any unusual patterns.



Perform an analysis of the variation in the "availability_365" column. Make sure to explore different "binwidth" values in your analysis.

```{r}
ggplot(data = listings) +
  geom_histogram(mapping = aes(x = availability_365), binwidth = 20)
ggplot(data = listings) + 
  geom_histogram(mapping = aes(x = availability_365), binwidth = 30)

```

* Which values are the most common? Why?
The most airbnbs have very low availability. This could because many airbnbs only rent out for a few days a year.

* Which values are rare? Why? Does that match your expectations?
The values are rare around 50, 150, 250 days. This could be because the number of days doesn't match the number of days in a multiples of months.

* Can you see any unusual patterns? What might explain them?

There is an unusal pattern because the number of airbnbs increasing at 75, 175, 275, and 365. This could again be because this amount of time could be more normal for people renting.


```{r}
listings
```


What seems to be the most common name (of a person) in the city you selected?

Marie 

```{r}
listings %>%
  count(host_name) %>%
  arrange(desc(n)) 
```

Do the number of review affect the price off the Airbnb? How? Why do you think this happens?

As the number of review increases the price increases up to 100 reviews, but then decreases as reviews increases past 100. This could be because slightly lower numbers of reviews could be positive and helpful for people trying to get feedback. The decrease in price after 200 reviews could be because there are more negative reviews or that it is so congested with reviews that it is not helpful. 

```{r}
smaller_price <- filter(listings, price < 500)
ggplot(data = smaller_price) + 
  geom_point(mapping = aes(x = number_of_reviews, y = price)) +
  geom_smooth(mapping = aes(x = number_of_reviews, y = price), se = FALSE)
```


What type of room tends to have the highest Airbnb price?

The hotel room airbnbs have the highest Airbnb price. 

```{r}
ggplot(data = smaller_price) +
  geom_boxplot(mapping = aes(x = reorder(room_type,price,FUN=median), y = price))
```

Suppose you could purchase a property in the city you selected, and that you could rent it to others as an Airbnb. In what neighborhood would you want to purchase your property? Why?

I would choose Elysée because it has the higheset average price out of all the neighborhoods, but also doesn't have m any Airbnbs in that area so there would be less competition.  Also Marie does not have many properties there. 

```{r}
ggplot(data = smaller_price) +
  geom_boxplot(mapping = aes(x = reorder(neighbourhood,price,FUN=median), y = price))
```

```{r}
Marie <- listings %>%
  filter(host_name == 'Marie')

Marie %>%
  count(neighbourhood) %>%
  arrange(desc(n))
```

Visit a real estate website and find a property that is for sale in the neighborhood you selected. Take notes of the price and address of the property.

Use your dataset to find what the average Airbnb price/night is in the neighborhood you selected.

Use your dataset to find what the average number of available nights per year is for an Airbnb in the neighborhood you selected.

Suppose you bought the property you selected above. If you were to rent it as an Airbnba t the average neighborhood price, for the average number of days, how long will it take you to break even?

```{r}
listings
```

```{r}
Elysee <- filter(listings, neighbourhood == "Élysée")

summarize(Elysee, avg_price = mean(price, na.rm = TRUE), avg_nights = mean(availability_365, na.rm = TRUE))

Elysee
```
https://www.residences-immobilier.com/en/75/annonce-vente-appartement-paris-8eme-2173583.html 

231$ a night, 161 nights a year

400,000 euros = 474,452$

231*161 = 37,191

474,452$/37,191 = 12.75

It would take about 13 years to pay remake the price of the apartment. 
















